---
title: "GWAS_2022_inference_result"
author: "Shufan Zhang"
date: "2023-10-12"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE,
  eval = TRUE,
  warning = FALSE, message = FALSE
)
```


```{r data input, include=FALSE}
setwd("/Users/lovely_shufan/Dropbox (Edison_Lab@UGA)/AMF/AMF Imaging 2022/0_inference_using_MaskRCNN_2021/2_infer_result/GA_GWAS_2022/")
getwd()
dir()
load("GWAS_2022_inference_data_wrangling_EDA.RData")
library(ggplot2)
library(viridis)
my_theme <- theme(text=element_text(size=10),
        plot.title = element_text(face = "bold",size = rel(1.2), hjust = 0.5),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        axis.ticks=element_blank(),
        axis.line=element_blank(),
        panel.border=element_blank(),
        legend.position="none",
        panel.grid.major=element_line(color='#eeeeee'))
```

## Exploratory Data Analysis of Inference Result

### Data cleaning
The best performing Mask RCNN model from 2021 was applied to ZEISS AxioScan7 images from the three untreated blocks in the Georgia GWAS 2022 experiment. Batch effect between training images and new AxioScan7 images are removed by centering the new images. Inference result was deduplicated. Segmentations with confidence score higher than 80%, size larger than 250 pixels are saved for the downstream analyses.

The image resolution of AxioScan 7 microscope is 0.120 micron^2/pixel. Segmentation areas were converted to mm^2.  

### Visualize model performance

```{r histograms of confidencescore, echo=FALSE, fig.retina = 1.5, fig.height=6, fig.width=9, fig.align = "center"}
#confidence score
library(ggplot2)
library(viridis)
ggplot(seg_clean_v3, aes(x=confidenceScore,fill=annotations))+
  geom_histogram(position = 'identity',binwidth = 0.005) +
  scale_fill_viridis(discrete=TRUE) +
  labs(fill="")+
  facet_grid(rows = vars(annotations), scales = "free_y")+
  theme_classic()+
  my_theme

ggplot(seg_clean_v3, aes(x=confidenceScore,fill=annotations))+
  geom_histogram(position = 'identity',binwidth = 0.005) +
  scale_fill_viridis(discrete=TRUE) +
  labs(fill="")+
  facet_grid(rows = vars(annotations), cols = vars(block), scales = "free_y")+
  theme_classic()+
  my_theme

ggplot(seg_clean_v3, aes(y=confidenceScore, x=log(area_mm2), color=annotations))+
  geom_point() +
  scale_color_viridis(discrete=TRUE) +
  labs(fill="")+
  facet_grid(rows = vars(annotations))+
  #ylim(c(0,25))+
  theme_classic()+
  my_theme

```

### Block effect
```{r exploratory analysis block effect count density, echo=FALSE}
spread(count_density_anno[,-c(4,5)], key = "annotations", value= countDensity, -)

pc.rin.arbc.rg <- lmer(percent_colonization^(1/3) ~ arb_count_scaled + region + (1|accession), data = col.overall.class.predictors.scaled, REML = TRUE)

cd_empty <- lmer(count_density^0.2 ~ arb_count_scaled + exH_count_scaled +
                      sp_count_scaled + ves_count_scaled + (1|accession), data = col.overall.class.predictors.scaled)

model1 <- lm(log(countDensity)~accession+block+annotations+annotations*block, data = count_density_blk_anno)
a <- aov(model1)
print(summary(a))
tukey_result <- TukeyHSD(a)
```

```{r visualize block effect count density, echo=FALSE, dpi=300, fig.width=7, fig.height=7, eval=FALSE}
#confidence score
library(ggplot2)
library(viridis)
library(dplyr)
factor(count_density_blk$accession , levels = cd_order) -> count_density_blk$accession
factor(count_density_blk$block , levels = c("Block2", "Block8", "Block10")) -> count_density_blk$block

factor(count_density_blk_anno$accession , levels = cd_order) -> count_density_blk_anno$accession
factor(count_density_blk_anno$block , levels = c("Block2", "Block8", "Block10")) -> count_density_blk_anno$block

count_density_blk_anno <- count_density_blk_anno %>%
  mutate(grouping = (match(accession, cd_order) - 1) %/% 50 + 1)
count_density_blk <- count_density_blk %>%
  mutate(grouping = (match(accession, cd_order) - 1) %/% 50 + 1)

ggplot(count_density_blk_anno, aes(x = accession, y = countDensity, color = block)) +
  geom_point() +
  labs(fill="")+
  facet_wrap(vars(grouping,annotations), scales = "free_x", nrow = 7)+
  scale_color_viridis(discrete=TRUE) +
  #ylim(c(0,25))+
  theme_classic()

ggplot(count_density_blk_anno, aes(x = block, y = countDensity)) +
  geom_boxplot() +
  labs(fill="")+
  scale_color_viridis(discrete=TRUE) +
  #ylim(c(0,25))+
  theme_classic()

ggplot(count_density_blk_anno, aes(x = block, y = percenCol, fill=annotations)) +
  geom_boxplot() +
  labs(fill="")+
  scale_color_viridis(discrete=TRUE) +
  #ylim(c(0,25))+
  theme_classic()

#ggplot(count_density_blk, aes(x = accession, y = countDensity, color = block)) +
#  geom_point() +
#  labs(fill="")+
#  facet_wrap(vars(grouping), scales = "free_x", nrow = 7)+
#  scale_color_viridis(discrete=TRUE) +
#  #ylim(c(0,25))+
#  theme_classic()+
#  my_theme

```

```{r exploratory analysis block effect percent col, echo=FALSE}
model2 <- lm(log(percentCol)~accession+block+annotations+annotations*block, data = percent_col_blk_anno)
a2 <- aov(model2)
print(summary(a2))
tukey_result <- TukeyHSD(a2)
```

```{r visualize block effect percent col, echo=FALSE, dpi=300, fig.width=7, fig.height=7, eval=FALSE}
#confidence score
library(ggplot2)
library(viridis)
library(dplyr)
factor(percent_col_blk$accession , levels = pc_order) -> percent_col_blk$accession
factor(percent_col_blk$block , levels = c("Block2", "Block8", "Block10")) -> percent_col_blk$block

factor(percent_col_blk_anno$accession , levels = cd_order) -> percent_col_blk_anno$accession
factor(percent_col_blk_anno$block , levels = c("Block2", "Block8", "Block10")) -> percent_col_blk_anno$block

percent_col_blk_anno <- percent_col_blk_anno %>%
  mutate(grouping = (match(accession, pc_order) - 1) %/% 50 + 1)
percent_col_blk_anno <- percent_col_blk_anno %>%
  mutate(grouping = (match(accession, pc_order) - 1) %/% 50 + 1)

ggplot(percent_col_blk_anno, aes(x = accession, y = percentCol, color = block)) +
  geom_point() +
  labs(fill="")+
  facet_wrap(vars(grouping,annotations), scales = "free_x", nrow = 7)+
  scale_color_viridis(discrete=TRUE) +
  #ylim(c(0,25))+
  theme_classic()+
  my_theme

#ggplot(count_density_blk, aes(x = accession, y = countDensity, color = block)) +
#  geom_point() +
#  labs(fill="")+
#  facet_wrap(vars(grouping), scales = "free_x", nrow = 7)+
#  scale_color_viridis(discrete=TRUE) +
#  #ylim(c(0,25))+
#  theme_classic()+
#  my_theme

```


### Correlation between percent col and count density
```{r correlation, echo=FALSE, dpi=300, fig.width=7, fig.height=7, eval=FALSE}
percent_col_blk_anno$key2 <- paste(percent_col_blk_anno$key,percent_col_blk_anno$annotations,sep = "_")
count_density_blk_anno$key2 <- paste(count_density_blk_anno$key,count_density_blk_anno$annotations,sep = "_")
merge(count_density_blk_anno, percent_col_blk_anno[,-c(1:6,8)], by = "key2") -> pheno_blk_anno

factor(pheno_blk_anno$accession , levels = pc_order) -> pheno_blk_anno$accession
factor(pheno_blk_anno$block , levels = c("Block2", "Block8", "Block10")) -> pheno_blk_anno$block

pheno_blk_anno <- pheno_blk_anno %>%
  mutate(grouping_pc = (match(accession, pc_order) - 1) %/% 50 + 1)

ggplot(pheno_blk_anno, aes(x = countDensity, y = percentCol, color = block)) +
  geom_point() +
  labs(fill="")+
  facet_wrap(vars(grouping,annotations), scales = "free_x", nrow = 7)+
  scale_color_viridis(discrete=TRUE) +
  #ylim(c(0,25))+
  theme_classic()+
  my_theme
```


## Data visualize
```{r overall percent colonization boxplots, dpi=300, fig.width=7, fig.height=7, eval=FALSE}
percent_col_anno$key <- paste(percent_col_anno$accession,percent_col_anno$annotations, sep = "_")
count_density_anno$key <- paste(count_density_anno$accession,count_density_anno$annotations, sep = "_")

merge(percent_col_anno, count_density_anno[,-c(1:5)], by = "key") -> pheno_anno
my_theme45 <- theme(text = element_text(size = 16),
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.title.y = element_text(margin = margin(r = 20), vjust = -0.25),
        legend.position = "right",legend.title= element_blank())

ggplot(pheno_anno, aes(fill=annotations, y=countDensity, x=factor(accession, levels = pc_order)))+ 
  geom_bar(stat="identity")+
  scale_fill_viridis_d()+
  theme_classic()+
  my_theme45+
  xlab("accession orderd in decreading percent colonization")+
  ylab("count density")
```



